Systematic Design of Optimal Low-Thrust Transfers for the Three-Body Problem \thanks{This research has been supported in part by NSF under the grants CMMI-1243000 (transferred from 1029551), CMMI-1335008, and CNS-1337722.}

### 2017-07-27 Response.tex

\EditTL{Let's consider putting a rough computation time. We may not need time for each iteration, but you may have some idea how much it did take for the whole optimization. We may add a simple statement like ``it took ?? hours under ??? CPU'' We just need to show that we didn't use any monster cluster over several weeks. }

@@

    Thank you for the comment, and we agree that there is significant repetition of material previously covered in other sources.
    The manuscript has been reduced and reorganized to enhance the clarity and clearly seperate previously material and the contributions of this paper.
    The introduction has been reduced and the repetition and discuss on the variational integrator has been removed.
    The second section has been retitiled to ``Problem Formulation and Mathematical Background'' and has subsections which cover all of the background material, namely the planar circular restricted three body problem, Jacobi integral, invariant manifolds and \Poincare maps, the variational integrator.
The next sections introduce the optimal control formulation for the reachability analysis and the numerical examples, followed by conclusions.
The section on variational integrators has been reduced and now only discusses the main developments of the integrator for the PCRTBP and omits the review of variational principles.

\item 
    \begin{itshape}
        Secondly, the contributions and findings of the authors feel highly preliminary and do not seem to evaluate their method at significant depth.
    \end{itshape}

    Thank you for the comment. 
    The numerical examples are meant to demonstrate the use of this approach and the feasibility of using the reachability set on a \Poincare section to determine transfer trajecotries.
    The examples give two demonstrations of finding a transfer trajectory in the PCRTBP and demonstrate combining several iterations of the reachability set for a transfer.

\item
    \begin{itshape}
While there are the beginnings of a good article within this manuscript, it is not sufficiently close to publishable form to be accepted.  I recommend that the paper be rejected and that the authors push their investigation a bit further.  The eventual publication should focus less on background and should provide more insight into the fundamental properties of their method, using more rigorously constructed and analyzed examples. \EditTL{The authors believe that the given two nontrivial numerical examples illustrate the efficacy of the proposed approach in designing low-thrust maneuvers through reachability set analysis on the Poincar\'{e} section. Furthermore, the manuscript has been completely reorganized to emphasize the contributions. In particular, Section 4 includes additional paragraphs and figures for (fill the blank)}
\end{itshape}

@@

\item 
    \begin{itshape}
I can sympathize that there was quite a long road to get to the point of understanding this topic and generating results in it. I hope that the authors will not find this verdict too discouraging, as I believe that stronger results are well within their reach.
\end{itshape}
\end{itemize}

SPECIFIC FEEDBACK:
\begin{enumerate}
    \item
    \begin{itshape}
*** Bloat
In the introduction, the exposition of the problem could be considerably condensed, given the venue of publication. Additionally, the somewhat sprawling discussion jumps back and forth, and many key points seem to be repeated more than necessary.  It needs to be organized into titled subsections with specific scope.
\end{itshape}

Thank you for the comment.
The introduction is reduced and many of the paragraphs have been reworded and combined. 
In addition, the extended discussion of the benefits of variational integrators is removed from the introduction.
The mathematical background has been condensed into a single section with seperate subsections which discuss some of the mathematical details.
The background section is focused on reviewing the relevant material rather than reintroducing topics.
In the interest of space the reorganized section is omitted in this response. \EditTL{In short, the manuscript is completely reorganized and all of the background materials are summarized in Section 2}.

@@

Thank you for the comment.
The reference has been noted in the discussion of reachability sets.
\begin{correction}
More recently, reachability theory has recently been applied to space systems~\cite{holzinger2009,komendera2012a,dellnitz2006}.
\end{correction}

\item
    \begin{itshape}
*** Variational Integrators
Many of the specific claims regarding the use of a variational integrator are unsubstantiated and, I believe, incorrect. Why is energy drift "even more detrimental" with regards to "the long term effects of low-thrust"?  I would need to be convinced that a non-geometric integrator's energy drift relative to a constant energy baseline is significantly different from its drift relative to a gradually shifting baseline corresponding to low thrust. The specific claims about the numerical stability of derivatives also don't seem related to energy preservation.

\end{itshape}

The main contribution is systematic design of low-thrust orbital maneuvers through reachability set analysis on the Poincar\'{e} section. The authors decided to utilize the variational integrator as it is more computationally efficient in simulating a complex dynamic system accurately.

More explicitly, while the structural stability of geometric numerical integrators is often illustrated by the conserved quantities such as energy or momentum, it has been shown that variational integrators are also effective in computing the rate of energy change for non-conservative systems accurately~[20]. This is particularly useful for low-thrust orbital maneuvers where the energy dissipate or accumulates slowly. Furthermore, it has been shown that the structure-preserving properties of geometric numerical integrators are also useful to numerical optimizations for improved computational efficiency and stability. For example, see the following reference.

\begin{itemize}
\item[-] Ober-Blöbaum, Sina, Oliver Junge, and Jerrold E. Marsden. "Discrete mechanics and optimal control: an analysis." ESAIM: Control, Optimisation and Calculus of Variations 17.2 (2011): 322-352.
\end{itemize}

In the revised manuscript, all of the comments for the variational integrator are re-examined and reorganized to Section 2.2, as also discussed in the response to the first reviewer.

@@

\item
    {\color{red}
    \begin{itshape}
Apart from these claims, I question the need for a variational integrator in the first place. The example in section 3.4 is not convincing because there is no apples-to-apples basis for the comparison; a variable-step integrator with unspecified tolerance is compared to a fixed-step integrator with an arbitrary step size.  Integrator comparisons should consider error in proportion to compute cost, which can be tuned by step size or tolerance.  Also, fixed step seems inappropriate for a path through diverse dynamic regimes, and it's not clear that the problem scope calls specifically for energy preservation (especially since the idea is to use low thrust to mitigate some time cost of exploiting the problem structure).  
\end{itshape}
}

In the revised manuscript, the computational properties of the variational integrator is explicitly compared with a general purpose integrator in~\Cref{fig:compare_energy}. According to the reviewer's suggestion, the error in energy conservation is illustrated for several computational costs. It is shown that the variational integrator exhibits lower conservation error  against a Runge-Kutta method for the similar level of CPU time.

With regards to the comments on the adaptive step size, the variational integrator can be generalized according to the multisymplectic framework, as in the following reference. 
\begin{itemize}
\item[-] Lew, Adrian, et al. "Asynchronous variational integrators." Archive for Rational Mechanics and Analysis 167.2 (2003): 85-146.
\end{itemize}
However, the above formulation for structure-preserving, adaptive time-step variational integrators is beyond the scope of this paper, and we use a fixed time step throughout the presented numerical results.

More importantly, it has been shown variational integrators are also effective in computing the rate of energy change for non-conservative systems accurately~[20], even for the fixed-step sizes.

@@

\item
    \begin{itshape}
Consider plotting the intersections of manifolds with the Poincare section as continuous curves.  
\end{itshape}

Thank you for the comment.
The manifold intersections with the \Poincare section are plotted as discrete points as the curves were generated in a discrete manner. 
A series of trajectories were generated by perturbing the trajectory at various points along the periodic orbit. 
As a result, we show that the intersection is a discrete approximation of the true intersection of the manifold with the section by using the discrete points rather than a continuous curve.

\item
    \begin{itshape}
Also consider zooming/scaling as necessary to show the details you discuss.
\end{itshape}

Thank you for the comment. The figure have been revised to adequately show the desired information in the revised manuscript. 
The figures are centered around the various intersections of the reachable set and the \Poincare section.
Furthermore, in the case of the much larger geostationary transfer requires it becomes more difficult to zoom in on the fine details.
However, a detailed view is presented which shows the various iterations of the reachability set computations.

@@

\item
    {\color{red}
    \begin{itshape}
*** Method/Examples
I urge a revisiting of the claims with regards to optimality.  The optimization method is being initialized with *something*, and stationary points are *local* extrema.  More precise claims and assurances about solution quality (and therefore the validity of the computed reachable set boundary) are needed.  
\end{itshape}
}

\EditTL{The presented numerical results satisfy the necessary conditions for optimality. As common in gradient-based numerical optimizations for non-convex systems, it guarantees local optimality. Applying various techniques to obtain the globally optimal solution is beyond the scope of this paper that is focused on systematic design of low-thrust orbital maneuvers thorough reachability set analysis on the Poincar\'{e} section.

To clarify these, the following comment is added at the first paragraph of Section 4.

\begin{correction}
The numerical examples presented in this section satisfy necessary conditions for local optimality, and obtaining a globally optimal solution is considered beyond the scope of this paper.
\end{correction}}

@@

\item
    \begin{itshape}
In the second example, what do you lose by chaining together multiple reachability-based steps?  Aren't you imposing that the trajectory routinely returns to the energy level of the Poincare section? How feasible would it be to conduct a single reachability analysis covering multiple revolutions? And can you use knowledge about directionality of movement on the Poincare section to limit the scope of your reachability search?
\end{itshape}

While the trajectories do return to the energy level of the \Poincare section, there is a transfer of ``energy'' between the states of the vehicle.
It is in this manner that the reachability set is enlarged to eventually include the target state.

Conducting a single reachability analysis becomes more complicated as the larger time of flight increases the sensitivity of the terminal states to variations in the initial costates. 
Furthermore, the larger time of flight creates a much larger and more chaotic reachable set for the analysis.
Through this work it was found that shorter trajectories tended to have more stable and consistent solutions.

\item
    {\color{red}
    \begin{itshape}
In general, please devote more depth to probing the strengths, weaknesses, and fundamental properties of the method.  Considering extreme cases can help with this.
\end{itshape}
}

\EditTL{Would you try to respond this. For the strengths, we can emphasize a systematic design of low-thrust orbital maneuvers, agains harecore optimizations. For the weakness, we can use several things that the first reviewer itemized, and also our previous response to that. }

\end{enumerate}
\bibliography{library}
\bibliographystyle{spmpsci}
\end{document}

### 2017-07-27 Response.tex

-Pros = boundedness in the energy behavior.  Very long term integration will not deviate from the known dynamical structure.

-Cons = you don't know where you are on the energy surface, just that you are on it somewhere.  This is still a problem in actual trajectory design.  Knowing the error is on an energy surface does not mean that you don't have error.
\end{itshape}

Thank you for the comment and the detail on the benefits/drawbacks of variational integrators. The authors agree that the general-purpose numerical integrators such as higher-order RK methods are suitable in many cases. The proposed approach utilizes the variational integrator, particularly due to its numerical structural stability of capturing long-term behaviors of small effects accurately and efficiently, which is critical in low-thrust orbital maneuvers.

While the structural stability of geometric numerical integrators is often illustrated by the conserved quantities such as energy or momentum, it has been shown that variational integrators are also effective in computing the rate of energy change for non-conservative systems~\cite{marsden2001}.

@@

To clarify these, additional discussion has been included in the paper in the subsection which introduces the variational integrator for the PCRTBP.
Much of the discussion of variational integrators in the introduction has also been removed.
The additional content, which discusses the relative benefits of variational integrators is copied below.

\begin{correction}
Both integration schemes result in trajectories that are initially nearly identical.
\Cref{fig:compare_energy} shows the mean Jacobi integral deviation over the entire simulation time as a function of computation time.
For a given computational effort, in the form of simulation run time, the variational integrator will provide a smaller energy deviation as compared to the conventional integration scheme.
Over long simulation horizons or with the addition of small control inputs the inability of conventional integration schemes to accurately track the system energy limits the applicability of conventional techniques in which energy conservation is mandatory for characterizing the solution space.
In spite of this improved energy behavior, the order and design of the variational integrator still play an important role in the accuracy of the state vector.
This work uses a second order integrator which has improved energy behavior but potentially lower state accuracy as compared to high order Runge Kutta methods. 
However, for this preliminary analysis the variational integrator provides reduced computational demands and is useful to characterize and explore the trade space in preliminary mission design scenarios.
In addition, it has been shown that variable step integrators, such as \texttt{ODE45} in Matlab, tend to degrade first-order gradient based methods, such as those used in~\cref{sec:optimal_control}~\cite{pellegrini2016}.
The fixed step size of the presented variational integrator mitigates the artificial ``noise'' of variable step integration schemes in gradient computation.
In summary, this paper utilizes the variational integrator to capture the long-term effects of low-thrust devices accurately and efficiently.
The structural stability during numerical integration is critical for numerical optimization in the subsequent developments.
\end{correction}
\end{itemize}
From here on out, comments specific to various sections of the paper follow:

@@

Thank you for the comment. Those numerical dissipations would depend on several parameters. But, for the particular numerical comparison illustrated by \Cref{fig:compare_energy}, such difference in conservation properties becomes noticeable in the simulation for 15 years. As suggested, the claims about the variational integrator have been qualified with some additional detail.
The correction is copied below.
\begin{correction}
As a result, these methods suffer from numerical instability and energy drift behaviors which make them ill-suited for the long-term propagation, e.g. on the order of decades to centuries, which is typical of the three-body problem. 
\end{correction}

\item
    \begin{itshape}
"Conventional integration techniques fail to capture the physical laws and geometric properties of the dynamic system [8]. As a result, the long term effects of low-thrust on the spacecraft trajectory are not accurately captured."  Again, I think situationally I would disagree.  Transfers on the order of 100's of days can maintain accuracies of 1e-10 (nondimensional units) in the Hamiltonian, which is generally considered a "small" amount of energy drift.
\end{itshape}

These sentences have been modified and combined to enhance clarity and reduce some of the unecessary bloat. 
In addition, some additional detail has been added to explicitly define the notion of ``long-term'' in the context of numerical integration. As shown by \Cref{fig:compare_energy}, the variational integrator exhibit smaller conservation errors than ODE45 at the similar computational efforts, for the particular numerical case considered. The modifications are copied below.
\begin{correction}
    Conventional Runge-Kutta integration techniques, such as those implemented in~\cite{mingotti2011,grebow2011}, fail to capture the physical laws and geometric properties of the dynamic system.
    As a result, these methods suffer from numerical instability and energy drift behaviors which make them ill-suited for the long-term propagation, e.g. on the order of decades to centuries, which is typical of the three-body problem. 
\end{correction}

@@

\begin{correction}
While the \Poincare~section is well defined for the planar case, as a 2D representation, it is more challenging to implement and visualize for the general 3D three-body problem.
\end{correction}

\item
    \begin{itshape}
-"This discrete update map, or variational integrator, shares the same geometric properties of the continuous time system and exhibits much better energy behavior than the traditional integration methods, especially over long transfer durations with small magnitude control inputs."  Yes, but the order of the variational integrator still matters.  The drift in the state vector accuracy is still important to a mission designer, even if it is guaranteed to lie on an energy surface.  I think you should indicate that this is more for preliminary analysis to characterize the transfers, when this drift in state vector accuracy is tolerable.  I.E., it's useful for the introductory phases of mission design when trying to globally characterize the design space.  Characterizing the scope of the analysis in this way would be helpful to the audience.
\end{itshape}

Thank you for the comment.
The section on the variational integrator has been reduced and this specific sentence has been removed. 
However, additional detail has been provided in a the variational integrator subsection which discusses this drawback.

\begin{correction}

Over long simulation horizons or with the addition of small control inputs the inability of conventional integration schemes to accurately track the system energy limits the applicability of conventional techniques in which energy conservation is mandatory for characterizing the solution space.
In spite of this improved energy behavior, the order and design of the variational integrator still play an important role in the accuracy of the state vector.
This work uses a second order integrator which has improved energy behavior but potentially lower state accuracy as compared to high order Runge Kutta methods. 
However, for this preliminary analysis the variational integrator provides reduced computational demands, as compared with general-purpose numerical integrators with a similar level of numerical error, and is useful to characterize and explore the trade space in preliminary mission design scenarios.

\end{correction}
\item 
    \begin{itshape}
"Optimal solutions are generally sensitive to small variations in the initial multipliers. As a result, the numerical stability of sensitivity derivatives is critical to accuracy and computational performance."  I don't agree that this implies that variational integrators are better for gradient-based methods.  The primary issue is that gradient-based methods that are solving optimal control problems are using first-order approximations to iterate on the problem.  The first-order approximation is what is poor, not necessarily the gradients.  Even if these gradients are of exact accuracy (and I agree that they should always be as accurate as possible), the first-order methods in the chaotic problem are what give the difficulties.  One could use a variational or standard integrator in a first-order gradient based method and nothing will change the fact that the method will be highly sensitive, save for partitioning the sensitivities as you did with your multiple shooting approach.  At best, you can change the statement to indicate that the accuracy in the gradients may improve, which is always helpful to gradient-based methods, but it's no guarantee that it eliminates major problem sensitivity.
\end{itshape}

@@

\begin{itemize}
\item[-] Lew, Adrian, et al. "Asynchronous variational integrators." Archive for Rational Mechanics and Analysis 167.2 (2003): 85-146.
\end{itemize}

However, the above formulation for structure-preserving, adaptive time-step variational integrators is beyond the scope of this paper, and we use a fixed time step throughout the presented numerical results.

Additional detail has been included in the manuscript discussing this potential drawback of variable step integrators.

\begin{correction}

In addition, it has been shown that variable step integrators, such as \texttt{ODE45} in Matlab, tend to degrade first-order gradient based methods, such as those used in~\cref{sec:optimal_control}~\cite{pellegrini2016}.
The presented results are based on the variational integrator with a fixed time step for numerical simulation and gradient computation.

\end{correction}

@@

\item 
    \begin{itshape}
"This results in an optimal solution rather than the suboptimal solution typical of direct optimal control methods." I disagree that suboptimality is typical of direct methods.  A direct method will typically solve the KKT conditions, which can be directly transcribed into the costates in the optimal control problem.  A direct method fully satisfying the KKT conditions can map to the indirect problem. If you want to strengthen this argument, then you should argue that it applies to direct methods which rely on coarse or suboptimal representations of the control.  For example, a hybrid direct method could use continuous optimal control theory control law, but directly minimize the cost function instead of solving the full TPBVP.  
\end{itshape}

Thank you for the comment.
The suboptimal claims of direct optimization has been reduced and modified.
Additional detail has been included on the merits of indirect optimal control vs. direct optimal control methods.
\begin{correction}
This results in an optimal solution rather than the suboptimal solution typical of direct optimal control methods which rely on coarse or suboptimal control parameterizations.
For example, a hybrid direct method may use a continuous control law developed from indirect optimal control theory while directly minimizing a cost function instead of solving the full two-point boundary value problem~\cite{ozimek2010a}.
\end{correction}

@@

Thank you for the comment.
The additional clarification has been added to the manuscript and copied below.
\begin{correction}
Over long simulation horizons or with the addition of small control inputs the inability of conventional integration schemes to accurately track the system energy limits the applicability of conventional techniques in which energy conservation is mandatory for characterizing the solution space.
\end{correction}

\item 
    \begin{itshape}
Section 5 - This method, while very useful in its given context is highly dependent on the use of the PCRTB in the framework of the current analysis.  This needs to somehow be conveyed.  3D transfers are significantly more challenging when it comes to exploiting Poincare sections.  It can be done, but more subplots, dimensions, and data storing are necessary to view the subspace - save to say, it is a more challenging endeavor that must be acknowledged for general purpose low-thrust mission design.
\end{itshape}

Thank you for the comment.
An additional paragraph has been added to the beginning of the mathematical background section which discusses our reachability based transfer approach.
This additional content is copied below.

\begin{correction}
The numerical examples presented in this section are designed in the context of the PCRTBP.
The dynamic environment has a four dimensional state space and offers a convient integration constant in the form of the Jacobi integral.
As a result, there are well defined methods to define and exploit \Poincare sections, which result in straightforward two-dimensional subspaces of the system.
Our approach uses the \Poincare section to approximate the reachability set on this reduced subspace.
As a result, this approach is more difficult to apply to three-dimensional transfers in the general three body problem.
\Poincare sections in the case of the general six dimensional state space are significantly more challenging and typically require more complicated visualization techniques. 
However, this is an area of active research and some of the authors future research is aimed at implementing this approach for non-planar transfer trajectories~\cite{kulumani2016d}.
\end{correction}

@@

    Thank you for the comment, and we agree that there is significant repetition of material previously covered in other sources.
    The manuscript has been reduced and reorganized to enhance the clarity and clearly seperate previously material and the contributions of this paper.
    The introduction has been reduced and the repetition and discuss on the variational integrator has been removed.
    The second section has been retitiled to ``Problem Formulation and Mathematical Background'' and has subsections which cover all of the background material, namely the planar circular restricted three body problem, Jacobi integral, invariant manifolds and \Poincare maps, the variational integrator.
The next sections introduce the optimal control formulation for the reachability analysis and the numerical examples, followed by conclusions.
The section on variational integrators has been reduced and now only discusses the main developments of the integrator for the PCRTBP and omits the review of variational principles.

\item 
    \begin{itshape}
        Secondly, the contributions and findings of the authors feel highly preliminary and do not seem to evaluate their method at significant depth.
    \end{itshape}

    Thank you for the comment. 
    The numerical examples are meant to demonstrate the use of this approach and the feasibility of using the reachability set on a \Poincare section to determine transfer trajecotries.
    The examples give two demonstrations of finding a transfer trajectory in the PCRTBP and demonstrate combining several iterations of the reachability set for a transfer.

\item
    \begin{itshape}
While there are the beginnings of a good article within this manuscript, it is not sufficiently close to publishable form to be accepted.  I recommend that the paper be rejected and that the authors push their investigation a bit further.  The eventual publication should focus less on background and should provide more insight into the fundamental properties of their method, using more rigorously constructed and analyzed examples. 
\end{itshape}

The authors believe that the given two nontrivial numerical examples illustrate the efficacy of the proposed approach in designing low-thrust maneuvers through reachability set analysis on the \Poincare section. 
Furthermore, the manuscript has been completely reorganized to emphasize the contributions and clearly seperate the background material. 
In particular,~\cref{sec:simulation} includes additional paragraphs and figures for the geostationary to periodic orbit transfer.
These additional figures demonstrate the ability to chain together multiple iterations of the reachability set computation to generate a nontrivial transfer trajectory.
Combined together, the numerical examples demonstrate the ability to depart from the vicinity of the Earth to the vicinity of the Moon using our proposed method of coupling low-thrust manuevers with reachability set analysis.

@@

\item 
    \begin{itshape}
I can sympathize that there was quite a long road to get to the point of understanding this topic and generating results in it. I hope that the authors will not find this verdict too discouraging, as I believe that stronger results are well within their reach.
\end{itshape}
\end{itemize}

SPECIFIC FEEDBACK:
\begin{enumerate}
    \item
    \begin{itshape}
*** Bloat
In the introduction, the exposition of the problem could be considerably condensed, given the venue of publication. Additionally, the somewhat sprawling discussion jumps back and forth, and many key points seem to be repeated more than necessary.  It needs to be organized into titled subsections with specific scope.
\end{itshape}

Thank you for the comment.
The introduction is reduced and many of the paragraphs have been reworded and combined. 
In addition, the extended discussion of the benefits of variational integrators is removed from the introduction.
The mathematical background has been condensed into a single section with seperate subsections which discuss some of the mathematical details.
The background section is focused on reviewing the relevant material rather than reintroducing topics.
In the interest of space the reorganized section is omitted in this response. 
In short, the manuscript is completely reorganized and all of the background materials are summarized in~\cref{sec:background}.

@@

Thank you for the comment.
The reference has been noted in the discussion of reachability sets.
\begin{correction}
More recently, reachability theory has recently been applied to space systems~\cite{holzinger2009,komendera2012a,dellnitz2006}.
\end{correction}

\item
    \begin{itshape}
*** Variational Integrators
Many of the specific claims regarding the use of a variational integrator are unsubstantiated and, I believe, incorrect. Why is energy drift "even more detrimental" with regards to "the long term effects of low-thrust"?  I would need to be convinced that a non-geometric integrator's energy drift relative to a constant energy baseline is significantly different from its drift relative to a gradually shifting baseline corresponding to low thrust. The specific claims about the numerical stability of derivatives also don't seem related to energy preservation.

\end{itshape}

The main contribution is systematic design of low-thrust orbital maneuvers through reachability set analysis on the \Poincare section. The authors decided to utilize the variational integrator as it is more computationally efficient in simulating a complex dynamic system accurately.

More explicitly, while the structural stability of geometric numerical integrators is often illustrated by the conserved quantities such as energy or momentum, it has been shown that variational integrators are also effective in computing the rate of energy change for non-conservative systems accurately~\cite{marsden2001}. This is particularly useful for low-thrust orbital maneuvers where the energy dissipate or accumulates slowly. Furthermore, it has been shown that the structure-preserving properties of geometric numerical integrators are also useful to numerical optimizations for improved computational efficiency and stability. For example, see the following reference.

@@

\item
    \begin{itshape}
Apart from these claims, I question the need for a variational integrator in the first place. The example in section 3.4 is not convincing because there is no apples-to-apples basis for the comparison; a variable-step integrator with unspecified tolerance is compared to a fixed-step integrator with an arbitrary step size.  Integrator comparisons should consider error in proportion to compute cost, which can be tuned by step size or tolerance.  Also, fixed step seems inappropriate for a path through diverse dynamic regimes, and it's not clear that the problem scope calls specifically for energy preservation (especially since the idea is to use low thrust to mitigate some time cost of exploiting the problem structure).  
\end{itshape}

In the revised manuscript, the computational properties of the variational integrator is explicitly compared with a general purpose integrator in~\Cref{fig:compare_energy}. According to the reviewer's suggestion, the error in energy conservation is illustrated for several computational costs. It is shown that the variational integrator exhibits lower conservation error  against a Runge-Kutta method for the similar level of CPU time.

With regards to the comments on the adaptive step size, the variational integrator can be generalized according to the multisymplectic framework, as in the following reference. 
\begin{itemize}
\item[-] Lew, Adrian, et al. "Asynchronous variational integrators." Archive for Rational Mechanics and Analysis 167.2 (2003): 85-146.
\end{itemize}
However, the above formulation for structure-preserving, adaptive time-step variational integrators is beyond the scope of this paper, and we use a fixed time step throughout the presented numerical results.

More importantly, it has been shown variational integrators are also effective in computing the rate of energy change for non-conservative systems accurately~\cite{marsden2001}, even for the fixed-step sizes.

@@

Thank you for the comment.
The manifold intersections with the \Poincare section are plotted as discrete points as the curves were generated in a discrete manner. 
A series of trajectories were generated by perturbing the trajectory at various points along the periodic orbit. 
As a result, we show that the intersection is a discrete approximation of the true intersection of the manifold with the section by using the discrete points rather than a continuous curve.

\item
    \begin{itshape}
Also consider zooming/scaling as necessary to show the details you discuss.
\end{itshape}

Thank you for the comment. The figure have been revised to adequately show the desired information in the revised manuscript. 
The figures are centered around the various intersections of the reachable set and the \Poincare section.
Furthermore, in the case of the much larger geostationary transfer requires it becomes more difficult to zoom in on the fine details.
However, a detailed view is presented which shows the various iterations of the reachability set computations.

\item
    \begin{itshape}
*** Method/Examples
I urge a revisiting of the claims with regards to optimality.  The optimization method is being initialized with *something*, and stationary points are *local* extrema.  More precise claims and assurances about solution quality (and therefore the validity of the computed reachable set boundary) are needed.  
\end{itshape}

The presented numerical results satisfy the necessary conditions for optimality. As common in gradient-based numerical optimizations for non-convex systems, it guarantees local optimality. Applying various techniques to obtain the globally optimal solution is beyond the scope of this paper that is focused on systematic design of low-thrust orbital maneuvers thorough reachability set analysis on the \Poincare section.

To clarify these, the following comment is added at the first paragraph of~\cref{sec:simulation}.

\begin{correction}
The numerical examples presented in this section satisfy necessary conditions for local optimality, and obtaining a globally optimal solution is considered beyond the scope of this paper.
\end{correction}

@@

\item
    \begin{itshape}
In the second example, what do you lose by chaining together multiple reachability-based steps?  Aren't you imposing that the trajectory routinely returns to the energy level of the Poincare section? How feasible would it be to conduct a single reachability analysis covering multiple revolutions? And can you use knowledge about directionality of movement on the Poincare section to limit the scope of your reachability search?
\end{itshape}

### 2017-07-27 Response.tex

While the trajectories do return to the energy level of the \Poincare section, there is a transfer of ``energy'' between the states of the vehicle.
It is in this manner that the reachability set is enlarged to eventually include the target state.

Conducting a single reachability analysis becomes more complicated as the larger time of flight increases the sensitivity of the terminal states to variations in the initial costates. 
Furthermore, the larger time of flight creates a much larger and more chaotic reachable set for the analysis.
Through this work it was found that shorter trajectories tended to have more stable and consistent solutions.

\item
    \begin{itshape}
In general, please devote more depth to probing the strengths, weaknesses, and fundamental properties of the method.  Considering extreme cases can help with this.
\end{itshape}

The authors believe that the given two nontrivial numerical examples illustrate the efficacy of the proposed approach in designing low-thrust maneuvers through reachability set analysis on the \Poincare section. 
The examples demonstrate the ability to generate transfer trajectories in the PCRTBP and can be easily extended to non-planar transfers~\cite{kulumani2016d}.
The proposed approach provides for a systematic design process for low-thrust orbital transfers, in contrast to other methods which must rely entirely on numerical optimization to determine trajectories.
This design process leverages the reachability set on a \Poincare section to greatly reduce the computational complexity of an unmotivated search of the reachable set.
Through the use of the reachability set, our proposed approach provides a methodical process to determine transfers, using single or multiple iterations.

There are several drawbacks of this approach.
First, it relies heavily on the PCRTBP, which limits the applicability to planar trajectories.
Furthermore, this reduces it's benefit to operational missions which are often non-planar, e.g. LEO to GTO missions.
Second, the method does not consider the effect of variable masss on the system dynamics. 
This is a focus of future work to incorporate the effect of variable mass on the optimal control problem.
Finally, the variational integrator provides for a bounded energy behavior but does not ensure exact energy over the simulation. 
As a result, there is still error in the variational integrator with respect to energy deviation.

The manuscript has been completely reorganized to emphasize the contributions and clearly seperate the background material. 
In particular,~\cref{sec:simulation} includes additional paragraphs and figures for the geostationary to periodic orbit transfer.
These additional figures demonstrate the ability to chain together multiple iterations of the reachability set computation to generate a nontrivial transfer trajectory.
Combined together, the numerical examples demonstrate the ability to depart from the vicinity of the Earth to the vicinity of the Moon using our proposed method of coupling low-thrust manuevers with reachability set analysis.

In addition, the conclusion has been modified to better demonstrate the strengths and weaknesses of our approach as well as avenues for future research.
The conclusion is copied below.

\begin{correction}

In this paper, an optimal transfer process which combines concepts of reachability and \Poincare~section is used to generate transfer between planar periodic orbits in the three-body problem.
The \Poincare~section allows for trajectory design on a lower dimensional phase space and simplifies the process.
The indirect optimal control formulation enables straightforward method of incorporating additional path and control constraints.
Direct optimal control techniques must rely entirely on the ability of the numerical optimzation routine to determine a feasible solution.
Our approach leverages the benefits of the reachability set, which encompass the maximum set of states achievable by the spacecraft over a fixed time horizon.
Using the reachability set on a \Poincare section reduces the dimensionality of the system and simplifies the analysis by avoiding the cost of a completely unmotivated exploration of the space.
However, the use of optimal control techniques leads to open loop trajectories that are not robust to model uncertainties or disturbances.
Furthermore, our approach relies heaviliy on the relative simplicity of the PCRTBP and this method is much more challenging for the non-planar case.

There is additional research to extend these results to more general transfer scenarios in future work.
The incorporation of fourth body perturbations, such as the Sun in the Earth-Moon system, offers an additional method of increasing the reachable set with the combined use of the solar perturbation and low-thrust propulsion.
In addition, the assumed acceleration magnitude is currently beyond the capabilities of current electric propulsion systems and future research is aimed at investigating smaller magnitude control inputs.
Furthermore, this analysis did not consider the effect of variable mass on the optimal control solution.
This will result in a more complicated optimal control problem and is a focus of future research.
Finally, Lyapunov control theory, which has previously been applied to the two-body problem, is being investigated in the hope of designing closed loop control schemes for this three-body scenario~\cite{chang2002}.
The addition of attitude dynamics and realistic pointing constraints would also significantly improve the applicability of this work.
\end{correction}

\end{enumerate}
% \bibliography{library}
% \bibliographystyle{spmpsci}
\end{document}
